### 2012-12-15 paper/paper.tex
filler crap on stemming do not push

\documentclass[preprint]{style}
\pagenumbering{arabic}

\usepackage{paralist}
\usepackage{graphicx}
\usepackage{url}
\usepackage{amsmath}
\usepackage{setspace}

% This gets rid of the block of whitespace on the first page
\makeatletter
\let\@copyrightspace\relax
\makeatother

\addtolength{\topmargin}{-.07in}
\addtolength{\textheight}{.07in}

%\addtolength{\oddsidemargin}{-0.08in}
%\addtolength{\evensidemargin}{-0.08in}
%\addtolength{\textwidth}{0.08in}
\setstretch{1.15}

\begin{document}

\title{Medical Concept Extraction}

\numberofauthors{3}
\author{
\alignauthor
Tristan Naumann\\
\affaddr{MIT EECS}
\email{tjn@mit.edu}
\alignauthor
Samantha Ainsley\\
\affaddr{MIT EECS}
\email{ainsley@mit.edu}
\alignauthor
Salman Ahmad\\
\affaddr{MIT EECS}
\email{saahmad@mit.edu}
}

\date{14 December 2012}

\maketitle
\begin{abstract}

This paper presents a system for extracting medical concepts from patient hospital records. Hospitals and medical practices are increasingly switching to electronic medical records--this trend has opened the doors to exciting applications for Natural Language Processing in medicine and biology. One such application is concept extraction: identifying high-level semantic labels in a body of text. Concept extraction on medical records is useful in automating analysis tasks that would otherwise have to be done manually. For example, hospital administrators may want to tabulate what medications are being prescribed the most. Insurance companies may want to ensure that patients receive proper care and are billed accordingly for fraud detection and auditing purposes. Government agencies may be interested in exploring and better understanding public health trends. Automatic concept extraction systems help to streamline such critical yet laborious tasks.

In service of that end, we present an algorithm and system that is capable of classifying words in medical records across four different medically-relevant categories: (1) problems, (2) tests, (3) treatments, and (4) non-medical terms. This is useful not only in granting end users the ability to analyze medical records, but also in supporting other higher-level Natural Language Processing and learning systems. This paper describes the system's design, implementation, and performance when tested on set of 500 pre-annotated medical records from Boston-area hospitals.

\end{abstract}

\section{Introduction}
TODO(SAM):
Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\section{Background}
TODO(SAM):
\subsection{Related Work}

\subsection{Machine Learning}

Talk about multi-class classification

\subsubsection{Support Vector Machines}

\subsubsection{Linear Regression}

\subsubsection{Conditional Random Fields}

\section{Challenges}

\subsection{High Dimensionality}

\subsection{Biased Data}

\subsection{Amount of Data}

Low amounts of train data with high amounts of tests data.

\subsection{Data Usage Agreements}
TODO(SAM):

\section{Design}

\subsection{Data Flow \& Classification Task}
Our system considers each document to be a stream of words grouped by sentences, and it attempts to provide the correct concept classification for each word. In order to identify the appropriate concept, our system considers natural language features for each document in successively broader contexts. Specifically, we first consider features using only the word itself as context, next we consider features which span the sentence, and last we consider n-gram features. In the sections that follow we discuss the linguistic features used in each context.

\subsection{Word Features}
Using only the word as context we are able to generate several interesting linguistic features. These features generally take on one of three forms: (1) intrinsic property of the word, (2) binary indicator for properties that we think are interesting, and (3) various techniques that map words into a small dimensional space such as stemming and word shape.

\subsubsection{Word}
The most obvious property of each word is perhaps the word itself. Since there is no linguistically meaningful, strict ordering over words, each word is represented as its own dimension. Consequently adding a baseline feature containing the word increases the dimension of our feature space to the size of the vocabulary in the training data. Nevertheless, the complete word serves as a fruitful feature in medical term classification in which many words carry the same meaning regardless of context. For example, terms such as "x-ray", "screen", "exam", "veal", and "sats" should be consistently recognized as test terms, just as "swelling", "failure", "trauma", and "syndrome" always indicate problems, and "therapy", "anesthesia", "vaccine", and "medication" are treatments. Granted, words like "heart" could be part of "heart rate" or "heart attack" and beg for more generic features.

\subsubsection{Length}
The next most obvious property of each word is its length. Unlike using the word as a feature, there is a strict ordering over length and consequently the addition of this feature increases the dimension of the space by only $1$. While the linguistic merit of length as a feature can be debated in conventional natural language processing, we felt it would play a large role in medical language in order to distinguish treatments, tests, and problems from words without a concept. Intuitively this is because often times treatments in the form of medications consist of long medical names, as do some of the words in medically described problems and tests.

\subsubsection{MIT Regular Expressions}
In order to capture some general characteristics of each word we included a set of regular expressions each of which adds a dimension to our classification. Such features help allow not only to pick up on traits within different medical categories--such as tests which are often abbreviated or contain numbers--but also to easily rule out non-medical terms such as days, phone numbers, etc. The characteristics we chose are as follows:
\begin{itemize}
\item {\tt INITCAP}: first letter is capitalized and alphabetic,
\item {\tt ALLCAPS}: entire word is capitalized and alphabetic,
\item {\tt CAPSMIX}: the word is a mix of capital and lowercase alphabetic characters,
\item {\tt HASDIGIT}: there is a digit in the word,
\item {\tt SINGLEDIGIT}: the token is a single digit,
\item {\tt DOUBLEDIGIT}: the token is a two digits,
\item {\tt FOURDIGITS}: the token is a series of four digits,
\item {\tt NATURALNUM}: the token is a natural number,
\item {\tt REALNUM}: the token is a real number,
\item {\tt ALPHANUM}: the token contains only digits and alphabetic characters,
\item {\tt HASDASH}: the token has a dash,
\item {\tt PUNCTUATION}: the token is punctuation,
\item {\tt PHONE1}: the token is a phone number,
\item {\tt PHONE2}: the token is a phone number with area code,
\item {\tt FIVEDIGIT}: the token is a series of five digits,
\item {\tt NOVOWELS}: the token does not contain vowels (i.e. abbreviations),
\item {\tt HASDASHNUMALPHA}: the token consists of a dash, numbers, and alphabetic characters, 
\item {\tt DATESEPERATOR}: the token is a date separator (e.g. {\tt [-/]})
\end{itemize}

\subsubsection{Porter Stemmer}
We already discussed the benefits of using entire words as features, but similar words come in many different forms. Motivated by generalization, we chose to use word stems as features and approached this feature extraction using three different stemming algorithms--two at the word-level and one at the sentence level (See 5.3.2). Namely, we used the Porter Stemming Algorithm, as developed by Martin Porter.

The stemmer is divided into a number of linear steps, five or six depending upon the definition of a step, that are used to produce the final stem. The following section will describe each step in turn, combined with a diagram displaying all the key steps of the algorithm. The following definitions regarding the stemmer need to be made before the steps can be explained. A consonant is a letter other than A, E, I, O, U and Y preceded by a consonant. For example the in the word boy the consonants are B and Y, but in try they are T and R. A vowel is any letter that is not a consonant. A list of consonants greater than or equal to length one will be denoted by a C and a similar list of vowels by a V.

Any word can therefore be represented by the single form; [C] (VC)m [V] Where the superscript m denotes m repetitions of VC and the square brackets [] denote the optional presence of their contents [19] The value m is called the measure of a word and can take any value greater than or equal to zero, and is used to decide whether a given suffix should be removed. All such rules are of the form; (condition) S1 -> S2 Which means that the suffix S1 is replaced by S2 if the remaining letters of S1 will satisfy the condition.

The first step of the algorithm is designed to deal with past participles and plurals. This step is the most complex and is separated into three parts in the original definition, 1a, 1b and 1c. The first part deals with plurals, for example sses -> ss and removal of s. The second part removes ed and ing, or performs eed ? ee where appropriate. The second part continues only if ed or ing is removed and transforms the remaining stem to ensure that certain suffices are recognised later. The third part simply transforms a terminal y to an i, this part is inserted as step 2 in the flowchart.

The remaining steps are relatively straightforward and contain rules to deal with different order classes of suffices, initially transforming double suffices to a single suffix and then removing suffices providing the relevant conditions are met.

Our CRF F1 scores went from 0.58 to 0.82

\subsubsection{Lancaster Stemmer}

The stemmer utilises a single table of rules, each of which may specify the removal or replacement of an ending. This technique of replacement is used to avoid the problem spelling exceptions as described earlier, by replacing endings rather than simply removing them the stemmer manages to do without a separate stage in the stemming process, i.e. no recoding or partial matching is required. This helps to maintain the efficiency of the algorithm, whilst still being effective. The rules are indexed by the last letter of the ending to allow efficient searching and are of the following form;

An ending of one or more characters, held in reverse order
An optional intact flag '*'
A digit specifying the removal total (zero or more)
An optional append string of one or more characters
A continuation symbol, '>' or '.'
The following example demonstrate how the rules are stored and used and how replacement can be used to negate the need for recoding. The rule nois4j> causes sion endings to be replaced by j. This acts as a pointer to the j section of the rules, leading to the following transformation; provision -> provij -> provid. The j-transformation is utilised to ensure that the terms provision and provide are correctly conflated to the stem provid.

The algorithm has four main steps detailed below, and presented in the flowchart 
1.Select relevant section; Inspect the final letter of the term and, if present, consider the first rule of the relevant section of the rule table.
2.Check applicability of rule; If final letters of term do not match rule, or intact settings are violated or acceptability conditions are not satisfied go to stage 4.
3.Apply Rule; Remove or reform ending as required and then check termination symbol, and either terminate or return to stage 1.
4.Look for another rule; Move to the next rule in table, if the section letter has changed then terminate, else go to stage 2.

\subsubsection{Word Shape}
In the same spirit of our word length and MIT RegEx features, we decided to look at the high-level shapes of individual words for better generalization. More specifically, we encode words into new strings that give a sense of their capitalization and punctuation patterns as well as approximate length. We tested a variety of open-sourced word shape classification methods including two methods written by Dan Klein, one of which simply classifies words as numeric, strictly upper or lower cased, or mixed, and a second, more fine-grained classifier that attempts to group words into equivalence classes by collapsing sequences of the same time. Additionally classifiers developed by Christopher Manning seek to make less distinctions mid-sequence by sorting words based on length. Each word shape classifier result added an additional dimension to our feature space. By naively including all word shape alternatives as features, we were able to improve our testbed CRF F1 score from 0.82 to 0.85. After further testing, we settled on \begin{verbatim}______\end{verbatim} as the most promising word shape metric because \begin{verbatim}______\end{verbatim}.
\begin{verbatim}
  ()_()
 ( ' ' )
   >.<
 (")_(")
\end{verbatim}

\subsubsection{Additional Feature Considerations}
The word-level features that we ultimately decided to include in system were chosen based on performance gain, but many others were explored in testing. These features were selected after close examination of our training data, and are worthy of discussion. It is possible, of course, that there is a more optimal feature set than our final choice that includes the features listed below:

\begin{itemize}

\item {\tt Metric Units}
A clear trend we noticed our medical record data was that numbers followed by metric weights (i.e. mg, milligrams, grams) are almost always part of treatment descriptions. Whereas numbers followed by metric lengths or areas are typically part of problem observations. We still believe this could be a useful feature for related systems.

\item {\tt Word Endings Indicative of Medical Problem Type}
Another obvious trend specific to medical problems are diseases ending in "itis" such as "bronchitis", "cellulites", "fibrositis", etc. and conditions ending in "ic" such as "diabetic", "anemic", etc. We added a regular expression to flag words with these endings, but the feature did little to nothing to improve performance. Such observations were likely already handled by our other word features.

\item {\tt Prognosis Locations}
Many diagnoses include locational information that should be tagged as part of the problem (i.e. "C4-C5 herniated disks"). We still think that a robust feature of this form could benefit our system, but our particular attempts proved unsuccessful.

\item {\tt Flag Words}
Certain terms such as those described in 5.2.1 are consistently labeled as "problems", "tests", or "treatments". We attempted to create small dictionaries of these terms and assign flags to words that fall into one of the three dictionaries. The issue with this strategy is that words that words that should be labeled as one of our three medical terms but do not appear in the dictionary are now more closely aligned with purely non-medical words. That said, this feature improved accuracy for our three categories, but also yielded far too many false positives.

\end{itemize}

\subsection{Sentence Features}
Even though the primary goal of our system is to tag concepts at the word level, these individual words often need to be considered in context. Before extracting the word-level features described in 5.2, we first analyze medical records at the sentence level and assign features to our words based on their usage within those sentences.

\subsubsection{Part of Speech}
A fruitful strategy for any concept extractor is to first tag words with labels indicating their Part of Speech in context. In our case, for example, whether an ambiguous term is used as an adjective versus a noun could distinguish it as part of a test rather than a problem. We used the NTLK max entropy Part of Speech Tagger, which is trained on the Penn TreeBank to tag individual sentences. We then iterate over the returned tags and add these as word-level features, adding another dimension to our feature set for each Part of Speech tag.

\subsubsection{WordNet Lemmatizer}
TODO(SAM):

\subsubsection{Formatted Text}
When reading through our training data, we noticed that test results are written in a variety of distinct forms that clearly separate these terms from problem or treatment terms. Since test result forms often spent multiple words such was "O2 sat -- 98\%" or "screen was positive". We wrote a series of regular expressions to capture these patterns with the test term as the first term in the sequence. We iterate over each word in the sentence and run our RegEx suite on the inclusive righthand sequence starting from that word. This feature adds only one additional dimension, and improved our CRF F1 test score from .85 to .90 by boosting our test term detection appreciably.

\subsection{N-gram Features}
TODO(SAM):

\subsubsection{Previous}

\subsubsection{Next}

\subsection{Parameter Tuning}

\begin{figure}
\begin{center}
	\includegraphics[width=1\columnwidth]{figures/parameter-selection.png}
\end{center}
\caption{An illustration of the grid search process to identify the best parameters values for training the SVM model.}
\label{fig:parameter_selection}
\end{figure}

\begin{itemize}

\item Grid search and grid search image

\item Include that PNG image from easy.py

\item Cross validation

\end{itemize}

\section{System}

\subsection{Code Architecture}

\begin{figure}
\begin{center}
	\includegraphics[width=1\columnwidth]{figures/system.png}
\end{center}
\caption{An overview of the system and codebase.}
\label{fig:system}
\end{figure}

We built a fairly comprehensive framework for applying machine learning techniques to NLP domains. We built out a series of useful and reusable library modules and command line utilities to perform various tasks. An overview of the codebase can be seen in Figure~\ref{fig:system} and  all of the code is available online at the following url:

{\tt https://github.com/tnaumann/ConceptExtraction}.

\subsubsection{Data Formats and File I/O}

\subsubsection{Feature Computation}

\subsubsection{Model Encapsulation and Serialization}

\subsubsection{Learning Algorithms}

\subsubsection{Data Organization and Directory Structure}

\subsubsection{Command-line Utilities}

\subsection{Web Service}

In addition to the various command line utilizes described in the previous section our system also includes a Web-based service that allows users to manually input a single sentence, select a model to use for prediction, and see the predicted labels that model generated for the input. Figure~\ref{fig:web_interface} shows a screenshot of the Web interface in use.

The Web service was written in Python and directly interoperates with the rest of the code base. To run the Web applications simply execute the {\tt web.py} file in the source tree. The Web app assumes that models have already been trained and placed into ``models'' directory in the source tree. By default, the application will host itself on port 5000. You can also see a live version of the Web service at:

{\tt http://saahmad.media.mit.edu:5000/ }

\begin{figure}
\begin{center}
	\includegraphics[width=1\columnwidth]{figures/web-interface.png}
\end{center}
\caption{The Web application interface that allows users to manually input a sentence and have it labeled using one of the previously-trained models. In this case we are using an SVM model that was trained using all of the features.}
\label{fig:web_interface}
\end{figure}

\subsection{Libraries}

Our code leveraged various 3rd party libraries for training the models and generating features. A summary of the libraries that we used is shown in the list below:

\begin{itemize}

\item {\tt libsvm} \cite{libsvm}: libsvm provides a fast implementation of Support Vector Machines. It includes a command line utility that takes a training data (with all of the features formatted in a sparse matrix representation) as input and outputs a SVM model trained on that data. It also includes a command line utility that allows users to specify a desired model and predict the classes on new input. Our code base forks process calls to these utilities to train SVM models, although a lot of the parameter tuning and feature selection is handled elsewhere in our codebase.

\item {\tt liblinear} \cite{liblinear}: lib linear is a linear classifier library that has special support multi-class classification. It also provides easy to use training and predicting command line utilities that accept a similar input format as libsvm.

\item {\tt crfsuite} \cite{crfsuite}: crfsuite is an implementation for conditional random fields and includes command line front ends to train a model on data and to predict the labels on an unseen data set. The data format for the input training files is also a sparse matrix representation but is slightly different from the format used by libsvm and lib linear. Our codebase nicely abstracts out these format differences and produces the right format depending on the model in use.

\item {\tt nltk} \cite{nltk}: nltk is a very popular NLP library written in Python. We leverage nltk to generate some of the features on our data. In particular, we utilize nltk's maximum-entropy part-of-speech tagging as well as many of its stemming algorithms.

\item {\tt flask} \cite{flask}: Flask is a micro Web framework written in Python. It contains a small yet useful set of features that allow developers to easily setup RESTful Web backends and an elegant tempting system to gracefully generate HTML. Flask is used to build the Web service portion of the project.

\end{itemize}

\section{Results}

\subsection{Data Set}

Talk about the competition and where the data came from. How large the data size is, how much we trained on, etc.

\subsection{Trained Models}

Describe the different feature-subsets that we trained and why.

\subsection{Runtime Performance}

How long each model took to train and how fast we can predict. Also discussion the hardware that was used to train and predict in terms of CPU speed, memory, etc.

\subsection{Classification Performance}

A table of a bunch of awesome results. We should present results by model type (SVM, CRF, LIN) and feature set.

\section{Discussion}

Talk about what we saw in terms of which was the best feature set and model.

Be sure to include examples of cool cases where it caught a difficult label

Be sure to include FAILURE cases

\section{Conclusion and Future Work}

\subsection{Dimension Collapse}

\subsection{Feature Selection}

\subsection{Parameter Selection}

\subsection{More Data}

\section{Acknowledgments}

We sincerely thank Dr. Robert Berwick  and Geza Kovacs
for their guidance, help, and support.

%%%% May the Flow (Max-Flow, that is) be with you all.

%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{citations}  % sigproc.bib is the name of the Bibliography in this case

\balancecolumns
% That's all folks!
\end{document}